---
title: "Data Analysis"
author: '659605'
date: "10/04/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(gam)
library(akima)
```

## R Markdown

```{r}
setwd("~/OneDrive/Uni/SOAS University of London/Modules/year 3/Machine Learning/Project/displacement_sudan/Code")
dataset <- read.csv("monthly_data_4.csv")
```
Selecting variables
```{r}
relevant_data <- dataset %>%
  select(mig.per.1000000, rain.dep.month, rain.month.before, rain.two.months.before,
         rain.sum.2.months, rain.sum.3.months, 
         temp.dep.month, temp.month.before, temp.two.months.before, 
         temp.sum.2.months, temp.sum.3.months,
         rel.rain.dep.month, rel.rain.month.before, rel.rain.two.months.before,
         rel.rain.sum.2.months, rel.rain.sum.3.months,
         rel.temp.dep.month, rel.temp.month.before, rel.temp.two.months.before,
         rel.temp.sum.2.months, rel.temp.sum.3.months, conflict_deaths,
         reason, dep.country)
```


dealing with missing values
```{r}
for (i in 1:nrow(relevant_data)) {
  if (is.na(relevant_data$mig.per.1000000[i]) == TRUE) relevant_data$mig.per.1000000[i] <- 0
}
```


Splitting data into training and testing data
```{r}
set.seed(1)
splits <- initial_split(relevant_data) #, strata = mig.per.1000000)
dataset_train <- training(splits)
dataset_test <- testing(splits)
```


Baseline model to compare more sophisticated models against.
```{r}
lm_model <- lm(mig.per.1000000 ~. -reason -dep.country, dataset_train)
tidy(lm_model)
```
Quadratic model 
```{r}
lm_model <- lm(mig.per.1000000 ~ (rel.rain.dep.month)^2 + temp.dep.month, dataset_train)
tidy(lm_model)
```

Figure 2a
```{r}
graph_data <- dataset %>%
  filter(admin2codes == "SS0303") %>%
  filter(month.dep != "2021-01-01")%>%
  data.frame() %>%
  mutate(month.dep = as.Date(month.dep))

p <- ggplot(graph_data, aes(x = month.dep)) +
  scale_x_date(date_labels = "%b")
p <- p + geom_bar(aes(y = rel.rain.dep.month, 
                      colour = "precipitation"),
                  stat = "identity",
                  fill = "#56B4E9")
p <- p + geom_line(aes(y = mig.per.1000000/7, 
                       colour = "Migration")) +
  geom_point(aes(y = mig.per.1000000/7, 
                       colour = "Migration"))
p <- p + scale_y_continuous(sec.axis = sec_axis(~.*7, name = "Migrants (per million)"))
p <- p + scale_colour_manual(values = c("#D55E00", "#56B4E9"))
p <- p + labs(y = "relative precipitation (in mm)",
              x = "Month of displacement in 2020", 
              colour = "Variables")

p <- p + theme(legend.position = c(0.1, 0.87), 
               legend.title = element_text(size = 8),
               legend.text = element_text(size = 8))
p <- p + theme(
    axis.title.y = element_text(color = "#0072B2", size=13),
    axis.title.y.right = element_text(color = "#D55E00", size=13)
  )
p <- p + ggtitle("Figure 2a: Precipitation compared to 10-year average \n and migration in Bor, South Sudan")
p 
```

Figure 2b
```{r}
graph_data <- dataset %>%
  filter(admin2codes == "SS0304") %>%
  filter(month.dep != "2021-01-01") %>%
  data.frame() %>%
  mutate(month.dep = as.Date(month.dep))

p <- ggplot(graph_data, aes(x = month.dep)) +
  scale_x_date(date_labels = "%b")
p <- p + geom_bar(aes(y = rel.rain.dep.month, 
                      colour = "precipitation"),
                  stat = "identity",
                  fill = "#56B4E9")
p <- p + geom_line(aes(y = mig.per.1000000, 
                       colour = "Migration")) +
  geom_point(aes(y = mig.per.1000000, 
                       colour = "Migration"))
p <- p + scale_y_continuous(sec.axis = sec_axis(~., name = "Migrants (per million)"))
p <- p + scale_colour_manual(values = c("#D55E00", "#56B4E9"))
p <- p + labs(y = "relative precipitation (in mm)",
              x = "Month of displacement in 2020", 
              colour = "Variables")

p <- p + theme(legend.position = c(0.1, 0.15), 
               legend.title = element_text(size = 8),
               legend.text = element_text(size = 8))
p <- p + theme(
    axis.title.y = element_text(color = "#0072B2", size=13),
    axis.title.y.right = element_text(color = "#D55E00", size=13)
  )
p <- p + ggtitle("Figure 2b: Precipitation compared to 10-year average \n and migration in Canal and Pigi, South Sudan")
p 
```


Fitting a GAM
```{r}
restricted <- dataset_train %>%
  filter(rel.rain.dep.month < 500) %>%
  rename(precipitation_difference = rel.rain.dep.month) %>%
  rename(temperature_difference = rel.temp.dep.month) %>%
  mutate(conflict_deaths = conflict_deaths/12) 
gam1 <- gam(mig.per.1000000 ~ s(precipitation_difference,3) + 
             s(temperature_difference,8) +
             s(conflict_deaths, 2), 
           data = restricted)
par(mfrow = c(1,3))
gam.plots <- plot(gam1, se = TRUE, col = "blue", 
     ylab = "Forced migration per 1 million")
```
Testing the performance of the GAM
```{r}
test.restricted <- dataset_test %>%
  rename(precipitation_difference = rel.rain.dep.month) %>%
  rename(temperature_difference = rel.temp.dep.month) %>%
  mutate(conflict_deaths = conflict_deaths/12)
gam.fit <- predict(gam1, newdata = test.restricted)
rmse <- mean(sqrt((gam.fit-dataset_test$mig.per.1000000)^2))
rmse
```

Checking GAM of absolute data (rather than relative) as above. It does not perform as well.
```{r}
restricted <- dataset_train %>%
  filter(rel.rain.dep.month < 550) %>%
  rename(precipitation = rain.dep.month) %>%
  rename(average_temperature = temp.dep.month) %>%
  mutate(conflict_deaths = conflict_deaths/12)
gam2 <- gam(mig.per.1000000 ~ s(precipitation,3) + 
             s(average_temperature,8) +
             s(conflict_deaths, 2), 
           data = restricted)
par(mfrow = c(1,3))
gam.plots <- plot(gam2, se = TRUE, col = "blue", 
     ylab = "Forced migration per 1 million per month")
```

```{r}
test.restricted <- dataset_test %>%
  rename(precipitation = rain.dep.month) %>%
  rename(average_temperature = temp.dep.month) %>%
  mutate(conflict_deaths = conflict_deaths/12)
gam.fit <- predict(gam2, newdata = test.restricted)
rmse <- mean(sqrt((gam.fit-dataset_test$mig.per.1000000)^2))
rmse
```



Fitting a GAM on the data from the month before the migration occured. This performs better.
```{r}
restricted <- dataset_train %>%
  filter(rel.rain.dep.month < 550) %>%
  rename(precipitation = rel.rain.month.before) %>%
  rename(average_temperature = rel.temp.month.before) %>%
  mutate(conflict_deaths = conflict_deaths/12)
gam3 <- gam(mig.per.1000000 ~ s(precipitation,3) + 
             s(average_temperature,8) +
             s(conflict_deaths, 2), 
           data = restricted)
test.restricted <- dataset_test %>%
  rename(precipitation = rel.rain.month.before) %>%
  rename(average_temperature = rel.temp.month.before) %>%
  mutate(conflict_deaths = conflict_deaths/12)
gam.fit <- predict(gam3, newdata = test.restricted)
rmse <- mean(sqrt((gam.fit-dataset_test$mig.per.1000000)^2))
rmse
```


Every variable is significant
```{r}
tidy(gam1)
```

The GAM passes the test against a baseline model
```{r}
rmse_baseline <- mean(sqrt((mean(dataset_test$mig.per.1000000)-dataset_test$mig.per.1000000)^2))
rmse_baseline
```

Figure 4
```{r, warning = FALSE}
setwd("~/OneDrive/Uni/SOAS University of London/Modules/year 3/Environment and Climate Crisis/AS1")
natural_disasters <- read.csv("number-of-natural-disaster-events.csv")
natural_disasters %>% 
  filter(Year > 1970) %>%
  filter(Entity == "Drought" | Entity == "Flood") %>%
  ggplot(aes(x = Year, y = Number.of.reported.natural.disasters..reported.disasters.)) +
  geom_line(aes(color = Entity)) +
  scale_color_manual(values = c("#D55E00", "#56B4E9")) + 
  ggtitle("Figure 4: Number of droughts and floods worldwide") +
  ylab("recorded cases") +
  xlab("Year")
```



